Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2011-04-06 faces_system.tex
Typesets properly without appendix, spills with appendix.

@@

\begin{algorithm}[thb]
\caption{\bf (Deformable Sparse Recovery and Classification for
Face Recognition)} \label{alg:deformable-src}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\y\in\Re^m$ and a deformation group $T$.
\STATE
{\bf for} each subject $i$,
\STATE \hspace{3mm} $\tau^{(0)}
\leftarrow I$.
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
\STATE \hspace{6mm}
$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
\STATE
\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$;
\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
residuals $\|\e\|_1$. \STATE Compute an average transformation
$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
$\ell^1$-minimization problem: \hspace{2mm}
$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
\end{small}
\end{algorithmic}
\end{algorithm}
%\vspace{-4mm}

@@

\subsection{System Implementation}
In this section, we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.

We have selected this algorithm because it strikes the best
balance between speed, accuracy, and scalability for our problem out of
many algorithms that we have tested. We refer the interested reader to
\cite{YangA2010-pp} for a more detailed discussion of competing
approaches.  On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory,
running on our database containing images size $80\times 60$
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm~\ref{alg:deformable-src} takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!

@@

We test the algorithms' ability to reject invalid images of the
88 subjects not appearing in the training database. As
mentioned before, the \emph{sparsity concentration index} (SCI)
is used as the outlier rejection rule. Given the sparse
representation $\x$ of a test image with respect to $K$
training classes, the SCI measures how concentrated the
coefficients are on a single class in the dataset and is
defined as in \cite{Wright2009-PAMI}:
\begin{displaymath}
\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
\|\x\|_1 - 1}{K - 1} \in [0,1] .
\end{displaymath}
It is easy to see that if $\textup{SCI}(\x) = 1$, the test
image is represented using images from one single subject
class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
evenly over all classes. Thus, we can choose a threshold $t \in
[0,1]$ for the proposed method and accept a test image as valid
if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
invalid. We compare this classifier to classifiers based on
thresholding the error residuals of NN, NS, LDA, and LBP.
\begin{figure}[t]
{\small
\centering
\begin{tabular}{@{}c@{}c@{}}
\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
(a) & \hspace{2mm}(b) \\
\end{tabular}
}
\vspace{-2mm}
\caption{{\bf ROC curves} for subject validation on Multi-PIE database,
(a) for all algorithms with iterative alignment, and
(b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=2.6in]{figures_pami/occ_block.png}
\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses. (b) The four overlapping blocks. (c) The sparse representation is calculated after alignment for each block independently. The red lines correspond to his true identity. (d) The true identity is successfully recovered by voting based on the SCI scores.}
\label{fig:occ-block}
\vspace{0mm}
\end{figure}

@@

\newcommand{\biospace}{\vspace{-2em}}

% biography section
\biospace
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/drew}}]{Andrew Wagner}
received his Bachelor's degree in General Engineering in 2003,
and his Master's degree in Electrical Engineering in 2006, from the University
of Illinois at Urbana-Champaign, where he is currently a Ph.D.
candidate in Electrical \& Computer Engineering.
His research interests include robotics, computer vision, and optimal control.
His recent work focuses on parallel algorithms for sparsity
based face recognition.\end{IEEEbiography}

@@

\biospace
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/arvind}}]{Zihan Zhou}
received... {\bf PLACEHOLDER}
\end{IEEEbiography}

@@

\biospace
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/yi}}]{Yi
Ma} received two Bachelors degree in Automation and Applied Mathematics
from  Tsinghua University, Beijing, China, in 1995. He received an Master
degree in Electrical Engineering and Computer Sciences (EECS) in 1997, a second
Master degree in Mathematics in 2000, and the Ph.D. degree in EECS in 2000, all
from the University of California at Berkeley. He is currently an associate
professor (with tenure) at the Department of Electrical and Computer
Engineering, University of Illinois at Urbana-Champaign, and since January 2009
has also served as research manager for the Visual Computing Group at Microsoft
Research Asia, Beijing, China.  \end{IEEEbiography}
\vfill

@@

>>>>

\usepackage{verbatim}

@@

\input{deformable-src.tex}

@@

\subsection{System Implementation}
The runtime of Algorithm~\ref{alg:deformable-src} is dominated
by the time spent solving two qualitatively similar $\ell_1$ minimization problems.
We have developed custom solvers for this purpose based on 
\emph{Augmented Lagrange Multiplier} (ALM) algorithm.  
We have selected this algorithm because it strikes the best
balance between speed, accuracy, and scalability for our problem out of
many algorithms that we have tested. We refer the reader to our
supplementary materials for a more in-depth discussion of our
solvers.  
For a more detailed discussion of competing
approaches, we refer the interested reader to \cite{YangA2010-pp} 
On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory,
running on our database containing images size $80\times 60$
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm~\ref{alg:deformable-src} takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!

@@

We test the algorithms' ability to reject invalid images of the
88 subjects not appearing in the training database. As
mentioned before, the \emph{sparsity concentration index} (SCI)
is used as the outlier rejection rule. Given the sparse
representation $\x$ of a test image with respect to $K$
training classes, the SCI measures how concentrated the
coefficients are on a single class in the dataset and is
defined as in \cite{Wright2009-PAMI}:
\begin{displaymath}
\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
\|\x\|_1 - 1}{K - 1} \in [0,1] .
\end{displaymath}
It is easy to see that if $\textup{SCI}(\x) = 1$, the test
image is represented using images from one single subject
class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
evenly over all classes. Thus, we can choose a threshold $t \in
[0,1]$ for the proposed method and accept a test image as valid
if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
invalid. We compare this classifier to classifiers based on
thresholding the error residuals of NN, NS, LDA, and LBP.
\begin{figure}[t]
{\small
\centering
\begin{tabular}{@{}c@{}c@{}}
\includegraphics[height = 1.6in]{figures_pami/pami_roc_revision2} & \hspace{2mm}
\includegraphics[height = 1.6in]{figures_pami/pami_roc2} \\
(a) & \hspace{2mm}(b) \\
\end{tabular}
}
\vspace{-2mm}
\caption{{\bf ROC curves} for subject validation on Multi-PIE database,
(a) for all algorithms with iterative alignment, and
(b) for the classical algorithms with manual alignment (indicated by a subscript ``m'').}\label{fig:roc-multipie}
\vspace{-2.5em}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=2.6in]{figures_pami/occ_block.png}
\caption{{\bf Using overlapping blocks to tackle contiguous occlusion.} (a) The test image, occluded by sunglasses. (b) The four overlapping blocks. (c) The sparse representation is calculated after alignment for each block independently. The red lines correspond to his true identity. (d) The true identity is successfully recovered by voting based on the SCI scores.}
\label{fig:occ-block}
\vspace{-3em}
\end{figure}

@@

% Biography section
\newcommand{\biospace}{\vspace{-4em}}
\biospace
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/drew}}]{Andrew Wagner}
received his Bachelor's degree in General Engineering in 2003,
and his Master's degree in Electrical Engineering in 2006, from the University
of Illinois at Urbana-Champaign, where he is currently a Ph.D.
candidate in Electrical \& Computer Engineering.
His research interests include robotics, computer vision, and optimal control.
His recent work focuses on parallel algorithms for sparsity
based face recognition.\end{IEEEbiography}

@@

\biospace
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/zihan}}]{Zihan Zhou}
Zihan Zhou received the bachelor's degree in automation from Tsinghua
University, China, in 2007. Since then, he has been with University of Illinois
at Urbana-Champaign, where he received the MS degree in electrical and computer
engineering and where he is now working toward the PhD degree. During the
summer of 2009, he was a research intern with Microsoft Research Asia, Beijing.
His research interests include computer vision, signal processing and machine
learning. 
\end{IEEEbiography}

@@

\biospace
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{bios/yi}}]{Yi
Ma} received two Bachelors degree in Automation and Applied Mathematics
from  Tsinghua University, Beijing, China, in 1995. He received an Master
degree in Electrical Engineering and Computer Sciences (EECS) in 1997, a second
Master degree in Mathematics in 2000, and the Ph.D. degree in EECS in 2000, all
from the University of California at Berkeley. He is currently an associate
professor (with tenure) at the Department of Electrical and Computer
Engineering, University of Illinois at Urbana-Champaign, and since January 2009
has also served as research manager for the Visual Computing Group at Microsoft
Research Asia, Beijing, China.  \end{IEEEbiography}

@@

\begin{comment}
\appendix[L1 Minimization via Augmented Lagrange Multiplier]
In this Appendix we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}, which is
repeated here for convenience. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.

We have selected this algorithm because it strikes the best
balance between speed, accuracy, and scalability for our problem out of
many algorithms that we have tested. We refer the interested reader to
\cite{YangA2010-pp} for a more detailed discussion of competing
approaches.  On a Mac Pro with
Dual-Core 2.66GHz Xeon processors and 4GB memory,
running on our database containing images size $80\times 60$
pixels from 109 subjects under 38 illuminations,
our C implementation of Algorithm~\ref{alg:deformable-src} takes
about 0.60 seconds per subject for alignment and about 2.0
seconds for global recognition. Compared to the highly
customized interior point method used in the conference version
of this paper \cite{Wagner2009-CVPR}, this new algorithm is
only slightly faster for per subject alignment. However, it is
much simpler to implement and it achieves a
\emph{speedup of more than a factor of 10} for global
recognition!
\end{comment}
